A general model for the evolution of nuptial gift-giving

### 2024-06-15 ms_biorxiv.Rmd
Make a biorxiv version for pre-print

---
title: "A general model for the evolution of nuptial gift-giving"
author: |
  | Anders P. Charmouh$^{1,4, 5}$, Trine Bilde$^{2}$, Greta Bocedi$^{1}$, A. Bradley Duthie$^{3}$
date: '[1] School of Biological Sciences, University of Aberdeen, UK. [2] Department of Biology, Aarhus University, Denmark. [3] Biological and Environmental Sciences, University of Stirling, UK.  [4] Bioinformatics Research Centre Aarhus University, University City 81, building 1872, 3rd floor, DK-8000 Aarhus C, Denmark [5] Email: apc@birc.au.dk'
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution-letters.csl
    reference_docx: docx_template.docx
bibliography: nuptial_refs.bib
header-includes:
- \usepackage{amsmath}
- \usepackage{natbib}
- \usepackage{lineno}
- \usepackage{caption}
- \usepackage[utf8]{inputenc}
- \bibliographystyle{amnatnat}
- \linenumbers
- \modulolinenumbers[2]
linestretch: 2
link-citations: yes
linkcolor: blue
csl: evolution-letters.csl
biblio-style: apalike
---

@@

Abstract
================================================================================

@@

Lay summary
================================================================================

If you have ever bought a gift to impress a prospective partner, you are not alone. 
Such behaviour is not unique to humans, but giving gifts during courtship is common in many different animals, including some species of insects, snails, birds, squid and spiders. 
In animals, such gifts can be costly, but offering them may be worth it from an evolutionary perspective if they increase the reproductive success of the individual who provides them. 
While using gifts during courtship (often called nuptial gifts) is widespread among animals, it is not universal. 
Why have some groups of animals evolved to engage in nuptial gift-giving while others have not? 
We explore this question by building a mathematical model, using computer simulations, and using data from a nuptial gift-giving species of spider to see how receiving nuptial gifts might affect reproductive success. 
We find conditions under which selection should favour males to provide nuptial gifts and females should exhibit a preference for gifts. 
These conditions explain why some species engage in nuptial gift-giving while others do not. Interestingly, we find that these inequalities differ between males and female, illustrating that the evolutionary interest of the two sexes do not always overlap.

Teaser text
================================================================================

The results presented provide a general explanation of the evolution of nuptial gift-giving which is a behaviour pattern that has evolved many times independently in diverse parts of the animal kingdom. 
The explanation is general but at the same time quantitatively precise, thus giving a rigorous answer to the conundrum of why nuptial gift-giving occurs in some groups of animals but not in others. 
Since research on nuptial gift-giving has, and continues to, inspire an enormous amount of work, the results presented here are highly relevant to an active field of both theoretical and empirical research.

@@

| Information                                                   | Page |
|---------------------------------------------------------------|------|
| S1: Description of the individual-based simulation model      |  27  |
| S2: Sensitivity analysis of mortality parameters in IBM       |  30  |
| S3: Alternative derivation of male fitness threshold          |  41  |
| S4: Operational sex ratio                                     |  43  |
| S5: Estimation of key model parameter using experimental data |  45  |
| S6: Separate evolution of male search and female choice       |  46  |

>>>>

---
title: "A general model for the evolution of nuptial gift-giving"
author: |
  | Anders P. Charmouh$^{1,4, 5}$, Trine Bilde$^{2}$, Greta Bocedi$^{1}$, A. Bradley Duthie$^{3}$
date: '[1] School of Biological Sciences, University of Aberdeen, UK. [2] Department of Biology, Aarhus University, Denmark. [3] Biological and Environmental Sciences, University of Stirling, UK.  [4] Bioinformatics Research Centre Aarhus University, University City 81, building 1872, 3rd floor, DK-8000 Aarhus C, Denmark [5] Email: apc@birc.au.dk'
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution-letters.csl
    reference_docx: docx_template.docx
bibliography: nuptial_refs.bib
header-includes:
- \usepackage{amsmath}
- \usepackage{natbib}
- \usepackage{lineno}
- \usepackage{caption}
- \usepackage[utf8]{inputenc}
- \bibliographystyle{amnatnat}
linestretch: 1
link-citations: yes
linkcolor: blue
csl: evolution-letters.csl
biblio-style: apalike
---

@@

********************************************************************************

@@

********************************************************************************

@@

| Information                                                   | Page |
|---------------------------------------------------------------|------|
| S1: Description of the individual-based simulation model      |  18  |
| S2: Sensitivity analysis of mortality parameters in IBM       |  20  |
| S3: Alternative derivation of male fitness threshold          |  29  |
| S4: Operational sex ratio                                     |  32  |
| S5: Estimation of key model parameter using experimental data |  33  |
| S6: Separate evolution of male search and female choice       |  34  |

### 2024-06-15 ms_biorxiv.tex
Make a biorxiv version for pre-print

\linenumbers
\modulolinenumbers[2]

@@

\setstretch{2}
\textbf{Key words:} Nuptial gifts, male search, mate interactions,
modelling, \emph{Pisaura}

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

@@

\hypertarget{lay-summary}{%
\section{Lay summary}\label{lay-summary}}

If you have ever bought a gift to impress a prospective partner, you are
not alone. Such behaviour is not unique to humans, but giving gifts
during courtship is common in many different animals, including some
species of insects, snails, birds, squid and spiders. In animals, such
gifts can be costly, but offering them may be worth it from an
evolutionary perspective if they increase the reproductive success of
the individual who provides them. While using gifts during courtship
(often called nuptial gifts) is widespread among animals, it is not
universal. Why have some groups of animals evolved to engage in nuptial
gift-giving while others have not? We explore this question by building
a mathematical model, using computer simulations, and using data from a
nuptial gift-giving species of spider to see how receiving nuptial gifts
might affect reproductive success. We find conditions under which
selection should favour males to provide nuptial gifts and females
should exhibit a preference for gifts. These conditions explain why some
species engage in nuptial gift-giving while others do not.
Interestingly, we find that these inequalities differ between males and
female, illustrating that the evolutionary interest of the two sexes do
not always overlap.

\hypertarget{teaser-text}{%
\section{Teaser text}\label{teaser-text}}

The results presented provide a general explanation of the evolution of
nuptial gift-giving which is a behaviour pattern that has evolved many
times independently in diverse parts of the animal kingdom. The
explanation is general but at the same time quantitatively precise, thus
giving a rigorous answer to the conundrum of why nuptial gift-giving
occurs in some groups of animals but not in others. Since research on
nuptial gift-giving has, and continues to, inspire an enormous amount of
work, the results presented here are highly relevant to an active field
of both theoretical and empirical research.

@@

\begin{figure}
\centering
\includegraphics{ms_refs_fixed_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{Fitness thresholds above which males increase their fitness by
searching for nuptial gifts (blue lines; Eq. 2) and females increase
their fitness by rejecting males that do not offer gifts (red lines; Eq.
3). Parameter space includes areas in which males do not search for
nuptial gifts and females are not choosy (A), males search but females
are not choosy (B), females would be choosy but males do not search (C),
and males search and females are choosy (D). Arrows in panel a indicate
the effect of increasing interaction rate (\(R\)), female time-out
(\(T_{\mathrm{f}}\)), and male search time (\(T_{\mathrm{m}}\)). As an
example, trajectories for \(T_{\mathrm{f}} = 2\), and
\(T_{\mathrm{m}} = \alpha\) are shown for values of \(R=2\) (panel a)
and \(R = 1\) (panel b). Females are assumed to be the choosy sex, which
is maintained as long as \(\alpha < T_{\mathrm{f}}\).}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_refs_fixed_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{The coevolution of male search and female choosiness as a
function of nuptial gift search time (\(\alpha\)). Points show where the
lower 95\% confidence interval of female choosiness (red) and male
search (blue) exceeds zero, indicating evolution of choosiness or
nuptial gift search. Each point includes data from 3200 replicate
simulations with identical starting conditions. Red and blue lines show
thresholds above which the mathematical model predicts that females
should be choosy and males should search, respectively (in agreement
with Figure 2). Minor deviations between the analytical model and the
simulation results are expected due to the finite population size, the
substantial stochasticity inherent to the simulation model, and the
potential for coevolution (for details, see Supporting Information S1).
The number of individuals in the population remained at or near carrying
capacity of \(K = 1000\). In each time step, up to 3000 total pair-wise
interactions occurred. Expected female processing time was set to
\(T_{\mathrm{f}}=2\) time steps, and \(\gamma\) and \(\alpha\) values in
the range {[}0.0, 2.0{]} and {[}0.1, 2.0{]}, respectively, were used.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_refs_fixed_files/figure-latex/unnamed-chunk-5-1.pdf}
\caption{The joint evolution of male search and female choosiness using
a nuptial gift fitness increment (\(\gamma\)) that was estimated from
experimental data for the species \emph{Pisaura mirabilis} (mean
\(\gamma = 3.29 \pm\) SE via propagation of error for estimates of
nuptial feeding a non-nuptial feeding groups: {[}1.47; 8.32{]}). Points
show where the 95\% confidence interval exceeds 0 for female choosiness
(red) and male search (blue). Each point includes data from 100
replicate simulations with identical starting conditions. The red line
shows the threshold above which females should be choosy and the blue
line shows the threshold above which males should search. As predicted
by the analytical model, both male search and female choosiness evolved
for a range of \(\gamma\) values around the empirical estimates. This
occurs because these values result in a \(\gamma\) above the fitness
threshold necessary for selection to favour male search of nuptial gifts
(blue line) and female choosiness (red line). The number of individuals
in the population remained at or near carrying capacity of \(K = 1000\).
In each time step, up to 3000 total pair-wise interactions occurred.
Female processing time was set to \(T_{\mathrm{f}}=2\) time steps, and
\(\gamma\) and \(\alpha\) values in the {[}0.2, 2.0{]}, respectively,
were used. The parameter \(\gamma\) was estimated as relative increase
in offspring production such that \(\gamma\) is the factor by which
fitness increases (relative to the baseline fitness) given a nuptial
gift (See Supporting Information S5).}
\end{figure}

@@

\begin{longtable}[]{@{}ll@{}}
\toprule
Information & Page \\
\midrule
\endhead
S1: Description of the individual-based simulation model & 27 \\
S2: Sensitivity analysis of mortality parameters in IBM & 30 \\
S3: Alternative derivation of male fitness threshold & 41 \\
S4: Operational sex ratio & 43 \\
S5: Estimation of key model parameter using experimental data & 45 \\
S6: Separate evolution of male search and female choice & 46 \\
\bottomrule
\end{longtable}

@@

\begin{figure}
\centering
\includegraphics{ms_refs_fixed_files/figure-latex/unnamed-chunk-6-1.pdf}
\caption{Figure S2.1: Evolution of both male search (blue) and female
choice (red) under different combinations of the mortality rates
\(\mu_{\mathrm{in}}\) and \(\mu_{\mathrm{out}}\) (mortality in time-in
and out, respectively). The y-axis is the threshold fitness that leads
to evolution of male search (blue) or female choice (red). The results
show noise, but no correlation between the value of the mortality
parameters and the propensity for male search and/or female choice to
evolve. For each of the \(25 \times 20\) combinations of
\(\mu_{\mathrm{in}}\) and \(\mu_{\mathrm{out}}\), 1600 replicate
simulations were run.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_refs_fixed_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{Figure S2.2 (\(T_{\mathrm{f}} = 0.5\)): The coevolution of male
search and female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
female choosiness (red) and male search (blue) exceeds zero, indicating
evolution of choosiness or nuptial gift search. Each point includes data
from approximately 1600 replicate simulations with identical starting
conditions (some parameter combinations have fewer replicates). Red and
blue lines show thresholds above which the mathematical model predicts
that females should be choosy and males should search, respectively. Up
to 3000 interactions occur between individuals in each time step
(\(\psi = 3\)), potentially resulting in a mating interaction. The
number of individuals in the population remained at or near carrying
capacity of \(K = 1000\). Expected female processing time was set to
\(T_{\mathrm{f}}=0.5\) time steps, and \(\gamma\) and \(\alpha\) values
in the range {[}0.5, 1.5{]} and {[}0.1, 2.1{]}, respectively, were
used.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_refs_fixed_files/figure-latex/unnamed-chunk-8-1.pdf}
\caption{Figure S2.3 (\(T_{\mathrm{f}} = 10.0\)): The coevolution of
male search and female choosiness as a function of nuptial gift search
time (\(\alpha\)). Points show where the lower 95\% confidence interval
of female choosiness (red) and male search (blue) exceeds zero,
indicating evolution of choosiness or nuptial gift search. Each point
includes data from approximately 1600 replicate simulations with
identical starting conditions (some parameter combinations include fewer
replicates). Red and blue lines show thresholds above which the
mathematical model predicts that females should be choosy and males
should search, respectively. Up to 3000 interactions occur between
individuals in each time step (\(\psi = 3\)), potentially resulting in a
mating interaction. The number of individuals in the population remained
at or near carrying capacity of \(K = 1000\). Expected female processing
time was set to \(T_{\mathrm{f}}=10.0\) time steps, and \(\gamma\) and
\(\alpha\) values in the range {[}0.5, 1.5{]} and {[}0.1, 2.1{]},
respectively, were used.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_refs_fixed_files/figure-latex/unnamed-chunk-9-1.pdf}
\caption{Figure S2.4 (\(\psi = 1\)): The coevolution of male search and
female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
where male search (blue) exceeds zero, indicating evolution of
choosiness or nuptial gift search. Each point includes data from
approximately 100 replicate simulations with identical starting
conditions (some parameter combinations had fewer replicates, minimum
400). Red and blue lines show thresholds above which the mathematical
model predicts that females should be choosy and males should search,
respectively. Up to 1000 interactions occur between individuals in each
time step (\(\psi = 1\)), potentially resulting in a mating interaction.
The number of individuals in the population remained at or near carrying
capacity of \(K = 1000\). Expected female processing time was set to
\(T_{\mathrm{f}}=2\) time steps, and \(\gamma\) and \(\alpha\) values in
the range {[}0.0, 1.4{]} and {[}0.1, 1.2{]}, respectively, were used.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_refs_fixed_files/figure-latex/unnamed-chunk-10-1.pdf}
\caption{Figure S2.5 (\(\psi = 2\)): The coevolution of male search and
female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
where male search (blue) exceeds zero, indicating evolution of
choosiness or nuptial gift search. Each point includes data from 100
replicate simulations with identical starting conditions (some parameter
combinations have fewer replicates). Red and blue lines show thresholds
above which the mathematical model predicts that females should be
choosy and males should search, respectively. Up to 2000 interactions
occur between individuals in each time step (\(\psi = 2\)), potentially
resulting in a mating interaction. The number of individuals in the
population remained at or near carrying capacity of \(K = 1000\).
Expected female processing time was set to \(T_{\mathrm{f}}=2\) time
steps, and \(\gamma\) and \(\alpha\) values in the range {[}0.0, 1.4{]}
and {[}0.1, 1.2{]}, respectively, were used.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_refs_fixed_files/figure-latex/unnamed-chunk-11-1.pdf}
\caption{Figure S2.6 (\(\psi = 4\)): The coevolution of male search and
female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
where male search (blue) exceeds zero, indicating evolution of
choosiness or nuptial gift search. Each point includes data from 100
replicate simulations with identical starting conditions (some parameter
combinations have fewer replicates). Red and blue lines show thresholds
above which the mathematical model predicts that females should be
choosy and males should search, respectively. Up to 4000 interactions
occur between individuals in each time step, potentially resulting in a
mating interaction (\(\psi = 4\)). The number of individuals in the
population remained at or near carrying capacity of \(K = 1000\).
Expected female processing time was set to \(T_{\mathrm{f}}=2\) time
steps, and \(\gamma\) and \(\alpha\) values in the range {[}0.0, 1.4{]}
and {[}0.1, 1.2{]}, respectively, were used.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_refs_fixed_files/figure-latex/unnamed-chunk-12-1.pdf}
\caption{Figure S2.7 (\(\psi = 6\)): The coevolution of male search and
female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
where male search (blue) exceeds zero, indicating evolution of
choosiness or nuptial gift search. Each point includes data from 100
replicate simulations with identical starting conditions (some parameter
combinations have fewer replicates). Red and blue lines show thresholds
above which the mathematical model predicts that females should be
choosy and males should search, respectively. Up to 6000 interactions
occur between individuals in each time step, potentially resulting in a
mating interaction (\(\psi = 6\)). The number of individuals in the
population remained at or near carrying capacity of \(K = 1000\).
Expected female processing time was set to \(T_{\mathrm{f}}=2\) time
steps, and \(\gamma\) and \(\alpha\) values in the range {[}0.0, 1.4{]}
and {[}0.1, 1.2{]}, respectively, were used.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_refs_fixed_files/figure-latex/unnamed-chunk-13-1.pdf}
\caption{Figure S2.8 (\(T_{\mathrm{m}} = 0.5\)): The coevolution of male
search and female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
where male search (blue) exceeds zero, indicating evolution of
choosiness or nuptial gift search. Each point includes data from 1600
replicate simulations with identical starting conditions. The red line
shows the threshold above which the mathematical model predicts that
females should be choosy (male thresholds for search are not shown
because simulations were initialised with males already searching for
nuptial gifts). Up to 3000 interactions occur between individuals in
each time step, potentially resulting in a mating interaction
(\(\psi = 3\)). The number of individuals in the population remained at
or near carrying capacity of \(K = 1000\). Expected female processing
time was set to \(T_{\mathrm{f}}=2\) time steps, and \(\gamma\) and
\(\alpha\) values in the range {[}0.0, 1.4{]} and {[}0.1, 1.2{]},
respectively, were used.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_refs_fixed_files/figure-latex/unnamed-chunk-14-1.pdf}
\caption{Figure S2.9 (\(T_{\mathrm{m}} = 2.5\)): The coevolution of male
search and female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
where male search (blue) exceeds zero, indicating evolution of
choosiness or nuptial gift search. Each point includes data from 1600
replicate simulations with identical starting conditions. The red line
shows the threshold above which the mathematical model predicts that
females should be choosy (male thresholds for search are not shown
because simulations were initialised with males already searching for
nuptial gifts). Up to 3000 interactions occur between individuals in
each time step, potentially resulting in a mating interaction
(\(\psi = 3\)). The number of individuals in the population remained at
or near carrying capacity of \(K = 1000\). Expected female processing
time was set to \(T_{\mathrm{f}}=2\) time steps, and \(\gamma\) and
\(\alpha\) values in the range {[}0.0, 1.4{]} and {[}0.1, 1.2{]},
respectively, were used.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_refs_fixed_files/figure-latex/unnamed-chunk-16-1.pdf}
\caption{Figure S6.1: The separate evolution of male search and female
choosiness as a function of nuptial gift search time. Points show where
the lower 95\% confidence interval of male search (blue) and female
choosiness (red) exceeds zero, indicating evolution of nuptial gift
search or choosiness. Each point includes data from \(2 \times 1600\)
replicate simulations with identical starting conditions. In the first
batch, male search was constant and initialized at
\(T_{\mathrm{m}} = \alpha\), and female choice was evolving. In the
second batch, male search was evolving, and there was no option for
female choice. The parameters \(T_{\mathrm{f}}=2\), and \(\gamma\) and
\(\alpha\) values were set within the range {[}0.1, 2.0{]} and {[}0.3,
1.7{]}, respectively.}
\end{figure}

>>>>

@@

\setstretch{1}
\textbf{Key words:} Nuptial gifts, male search, mate interactions,
modelling, \emph{Pisaura}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

@@

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

@@

\begin{figure}
\centering
\includegraphics{ms_biorxiv_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{Fitness thresholds above which males increase their fitness by
searching for nuptial gifts (blue lines; Eq. 2) and females increase
their fitness by rejecting males that do not offer gifts (red lines; Eq.
3). Parameter space includes areas in which males do not search for
nuptial gifts and females are not choosy (A), males search but females
are not choosy (B), females would be choosy but males do not search (C),
and males search and females are choosy (D). Arrows in panel a indicate
the effect of increasing interaction rate (\(R\)), female time-out
(\(T_{\mathrm{f}}\)), and male search time (\(T_{\mathrm{m}}\)). As an
example, trajectories for \(T_{\mathrm{f}} = 2\), and
\(T_{\mathrm{m}} = \alpha\) are shown for values of \(R=2\) (panel a)
and \(R = 1\) (panel b). Females are assumed to be the choosy sex, which
is maintained as long as \(\alpha < T_{\mathrm{f}}\).}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_biorxiv_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{The coevolution of male search and female choosiness as a
function of nuptial gift search time (\(\alpha\)). Points show where the
lower 95\% confidence interval of female choosiness (red) and male
search (blue) exceeds zero, indicating evolution of choosiness or
nuptial gift search. Each point includes data from 3200 replicate
simulations with identical starting conditions. Red and blue lines show
thresholds above which the mathematical model predicts that females
should be choosy and males should search, respectively (in agreement
with Figure 2). Minor deviations between the analytical model and the
simulation results are expected due to the finite population size, the
substantial stochasticity inherent to the simulation model, and the
potential for coevolution (for details, see Supporting Information S1).
The number of individuals in the population remained at or near carrying
capacity of \(K = 1000\). In each time step, up to 3000 total pair-wise
interactions occurred. Expected female processing time was set to
\(T_{\mathrm{f}}=2\) time steps, and \(\gamma\) and \(\alpha\) values in
the range {[}0.0, 2.0{]} and {[}0.1, 2.0{]}, respectively, were used.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_biorxiv_files/figure-latex/unnamed-chunk-5-1.pdf}
\caption{The joint evolution of male search and female choosiness using
a nuptial gift fitness increment (\(\gamma\)) that was estimated from
experimental data for the species \emph{Pisaura mirabilis} (mean
\(\gamma = 3.29 \pm\) SE via propagation of error for estimates of
nuptial feeding a non-nuptial feeding groups: {[}1.47; 8.32{]}). Points
show where the 95\% confidence interval exceeds 0 for female choosiness
(red) and male search (blue). Each point includes data from 100
replicate simulations with identical starting conditions. The red line
shows the threshold above which females should be choosy and the blue
line shows the threshold above which males should search. As predicted
by the analytical model, both male search and female choosiness evolved
for a range of \(\gamma\) values around the empirical estimates. This
occurs because these values result in a \(\gamma\) above the fitness
threshold necessary for selection to favour male search of nuptial gifts
(blue line) and female choosiness (red line). The number of individuals
in the population remained at or near carrying capacity of \(K = 1000\).
In each time step, up to 3000 total pair-wise interactions occurred.
Female processing time was set to \(T_{\mathrm{f}}=2\) time steps, and
\(\gamma\) and \(\alpha\) values in the {[}0.2, 2.0{]}, respectively,
were used. The parameter \(\gamma\) was estimated as relative increase
in offspring production such that \(\gamma\) is the factor by which
fitness increases (relative to the baseline fitness) given a nuptial
gift (See Supporting Information S5).}
\end{figure}

@@

\begin{longtable}[]{@{}ll@{}}
\toprule
Information & Page \\
\midrule
\endhead
S1: Description of the individual-based simulation model & 18 \\
S2: Sensitivity analysis of mortality parameters in IBM & 20 \\
S3: Alternative derivation of male fitness threshold & 29 \\
S4: Operational sex ratio & 32 \\
S5: Estimation of key model parameter using experimental data & 33 \\
S6: Separate evolution of male search and female choice & 34 \\
\bottomrule
\end{longtable}

@@

\begin{figure}
\centering
\includegraphics{ms_biorxiv_files/figure-latex/unnamed-chunk-6-1.pdf}
\caption{Figure S2.1: Evolution of both male search (blue) and female
choice (red) under different combinations of the mortality rates
\(\mu_{\mathrm{in}}\) and \(\mu_{\mathrm{out}}\) (mortality in time-in
and out, respectively). The y-axis is the threshold fitness that leads
to evolution of male search (blue) or female choice (red). The results
show noise, but no correlation between the value of the mortality
parameters and the propensity for male search and/or female choice to
evolve. For each of the \(25 \times 20\) combinations of
\(\mu_{\mathrm{in}}\) and \(\mu_{\mathrm{out}}\), 1600 replicate
simulations were run.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_biorxiv_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{Figure S2.2 (\(T_{\mathrm{f}} = 0.5\)): The coevolution of male
search and female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
female choosiness (red) and male search (blue) exceeds zero, indicating
evolution of choosiness or nuptial gift search. Each point includes data
from approximately 1600 replicate simulations with identical starting
conditions (some parameter combinations have fewer replicates). Red and
blue lines show thresholds above which the mathematical model predicts
that females should be choosy and males should search, respectively. Up
to 3000 interactions occur between individuals in each time step
(\(\psi = 3\)), potentially resulting in a mating interaction. The
number of individuals in the population remained at or near carrying
capacity of \(K = 1000\). Expected female processing time was set to
\(T_{\mathrm{f}}=0.5\) time steps, and \(\gamma\) and \(\alpha\) values
in the range {[}0.5, 1.5{]} and {[}0.1, 2.1{]}, respectively, were
used.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_biorxiv_files/figure-latex/unnamed-chunk-8-1.pdf}
\caption{Figure S2.3 (\(T_{\mathrm{f}} = 10.0\)): The coevolution of
male search and female choosiness as a function of nuptial gift search
time (\(\alpha\)). Points show where the lower 95\% confidence interval
of female choosiness (red) and male search (blue) exceeds zero,
indicating evolution of choosiness or nuptial gift search. Each point
includes data from approximately 1600 replicate simulations with
identical starting conditions (some parameter combinations include fewer
replicates). Red and blue lines show thresholds above which the
mathematical model predicts that females should be choosy and males
should search, respectively. Up to 3000 interactions occur between
individuals in each time step (\(\psi = 3\)), potentially resulting in a
mating interaction. The number of individuals in the population remained
at or near carrying capacity of \(K = 1000\). Expected female processing
time was set to \(T_{\mathrm{f}}=10.0\) time steps, and \(\gamma\) and
\(\alpha\) values in the range {[}0.5, 1.5{]} and {[}0.1, 2.1{]},
respectively, were used.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_biorxiv_files/figure-latex/unnamed-chunk-9-1.pdf}
\caption{Figure S2.4 (\(\psi = 1\)): The coevolution of male search and
female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
where male search (blue) exceeds zero, indicating evolution of
choosiness or nuptial gift search. Each point includes data from
approximately 100 replicate simulations with identical starting
conditions (some parameter combinations had fewer replicates, minimum
400). Red and blue lines show thresholds above which the mathematical
model predicts that females should be choosy and males should search,
respectively. Up to 1000 interactions occur between individuals in each
time step (\(\psi = 1\)), potentially resulting in a mating interaction.
The number of individuals in the population remained at or near carrying
capacity of \(K = 1000\). Expected female processing time was set to
\(T_{\mathrm{f}}=2\) time steps, and \(\gamma\) and \(\alpha\) values in
the range {[}0.0, 1.4{]} and {[}0.1, 1.2{]}, respectively, were used.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_biorxiv_files/figure-latex/unnamed-chunk-10-1.pdf}
\caption{Figure S2.5 (\(\psi = 2\)): The coevolution of male search and
female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
where male search (blue) exceeds zero, indicating evolution of
choosiness or nuptial gift search. Each point includes data from 100
replicate simulations with identical starting conditions (some parameter
combinations have fewer replicates). Red and blue lines show thresholds
above which the mathematical model predicts that females should be
choosy and males should search, respectively. Up to 2000 interactions
occur between individuals in each time step (\(\psi = 2\)), potentially
resulting in a mating interaction. The number of individuals in the
population remained at or near carrying capacity of \(K = 1000\).
Expected female processing time was set to \(T_{\mathrm{f}}=2\) time
steps, and \(\gamma\) and \(\alpha\) values in the range {[}0.0, 1.4{]}
and {[}0.1, 1.2{]}, respectively, were used.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_biorxiv_files/figure-latex/unnamed-chunk-11-1.pdf}
\caption{Figure S2.6 (\(\psi = 4\)): The coevolution of male search and
female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
where male search (blue) exceeds zero, indicating evolution of
choosiness or nuptial gift search. Each point includes data from 100
replicate simulations with identical starting conditions (some parameter
combinations have fewer replicates). Red and blue lines show thresholds
above which the mathematical model predicts that females should be
choosy and males should search, respectively. Up to 4000 interactions
occur between individuals in each time step, potentially resulting in a
mating interaction (\(\psi = 4\)). The number of individuals in the
population remained at or near carrying capacity of \(K = 1000\).
Expected female processing time was set to \(T_{\mathrm{f}}=2\) time
steps, and \(\gamma\) and \(\alpha\) values in the range {[}0.0, 1.4{]}
and {[}0.1, 1.2{]}, respectively, were used.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_biorxiv_files/figure-latex/unnamed-chunk-12-1.pdf}
\caption{Figure S2.7 (\(\psi = 6\)): The coevolution of male search and
female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
where male search (blue) exceeds zero, indicating evolution of
choosiness or nuptial gift search. Each point includes data from 100
replicate simulations with identical starting conditions (some parameter
combinations have fewer replicates). Red and blue lines show thresholds
above which the mathematical model predicts that females should be
choosy and males should search, respectively. Up to 6000 interactions
occur between individuals in each time step, potentially resulting in a
mating interaction (\(\psi = 6\)). The number of individuals in the
population remained at or near carrying capacity of \(K = 1000\).
Expected female processing time was set to \(T_{\mathrm{f}}=2\) time
steps, and \(\gamma\) and \(\alpha\) values in the range {[}0.0, 1.4{]}
and {[}0.1, 1.2{]}, respectively, were used.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_biorxiv_files/figure-latex/unnamed-chunk-13-1.pdf}
\caption{Figure S2.8 (\(T_{\mathrm{m}} = 0.5\)): The coevolution of male
search and female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
where male search (blue) exceeds zero, indicating evolution of
choosiness or nuptial gift search. Each point includes data from 1600
replicate simulations with identical starting conditions. The red line
shows the threshold above which the mathematical model predicts that
females should be choosy (male thresholds for search are not shown
because simulations were initialised with males already searching for
nuptial gifts). Up to 3000 interactions occur between individuals in
each time step, potentially resulting in a mating interaction
(\(\psi = 3\)). The number of individuals in the population remained at
or near carrying capacity of \(K = 1000\). Expected female processing
time was set to \(T_{\mathrm{f}}=2\) time steps, and \(\gamma\) and
\(\alpha\) values in the range {[}0.0, 1.4{]} and {[}0.1, 1.2{]},
respectively, were used.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_biorxiv_files/figure-latex/unnamed-chunk-14-1.pdf}
\caption{Figure S2.9 (\(T_{\mathrm{m}} = 2.5\)): The coevolution of male
search and female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
where male search (blue) exceeds zero, indicating evolution of
choosiness or nuptial gift search. Each point includes data from 1600
replicate simulations with identical starting conditions. The red line
shows the threshold above which the mathematical model predicts that
females should be choosy (male thresholds for search are not shown
because simulations were initialised with males already searching for
nuptial gifts). Up to 3000 interactions occur between individuals in
each time step, potentially resulting in a mating interaction
(\(\psi = 3\)). The number of individuals in the population remained at
or near carrying capacity of \(K = 1000\). Expected female processing
time was set to \(T_{\mathrm{f}}=2\) time steps, and \(\gamma\) and
\(\alpha\) values in the range {[}0.0, 1.4{]} and {[}0.1, 1.2{]},
respectively, were used.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{ms_biorxiv_files/figure-latex/unnamed-chunk-16-1.pdf}
\caption{Figure S6.1: The separate evolution of male search and female
choosiness as a function of nuptial gift search time. Points show where
the lower 95\% confidence interval of male search (blue) and female
choosiness (red) exceeds zero, indicating evolution of nuptial gift
search or choosiness. Each point includes data from \(2 \times 1600\)
replicate simulations with identical starting conditions. In the first
batch, male search was constant and initialized at
\(T_{\mathrm{m}} = \alpha\), and female choice was evolving. In the
second batch, male search was evolving, and there was no option for
female choice. The parameters \(T_{\mathrm{f}}=2\), and \(\gamma\) and
\(\alpha\) values were set within the range {[}0.1, 2.0{]} and {[}0.3,
1.7{]}, respectively.}
\end{figure}
